---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(seminr)
library(parallel)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# SEMinR

The `seminr` package provides a natural syntax for researchers to describe PLS structural equation models.

## Documentation

The vignette for Seminr can be found in the [seminr/inst/doc/](https://github.com/ISS-Analytics/seminr/blob/master/inst/doc/SEMinR.html) folder or by running the `vignette("SEMinR")` command after installation.

Demo code for use of Seminr can be found in the [seminr/demo/](https://github.com/ISS-Analytics/seminr/tree/master/demo) folder or by running the `demo("seminr-contained")`, `demo("seminr-ecsi")` or `demo("seminr-interaction")` commands after installation.

## Installation

You can install seminr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ISS-Analytics/seminr")
```

## Usage

Seminr can be used to create a plsm model, to estimate the model and to perform bootstrapping.

```{r example}
# seminr syntax for creating measurement model
mobi_mm <- measure(
  reflective("Image",        multi_items("IMAG", 1:5)),
  reflective("Expectation",  multi_items("CUEX", 1:3)),
  reflective("Value",        multi_items("PERV", 1:2)),
  reflective("Satisfaction", multi_items("CUSA", 1:3))
)

# interaction factors must be created after the measurement model is defined
mobi_xm <- interact(
  interaction_combo("Image", "Expectation"),
  interaction_combo("Image", "Value")
)

# structural model: note that name of the interactions factor should be
#  the names of its two main factors joined by a '.' in between.
mobi_sm <- structure(
  paths(to = "Satisfaction",
        from = c("Image", "Expectation", "Value",
                 "Image.Expectation", "Image.Value"))
)

# Load data, assemble model, and estimate using semPLS
data("mobi", package = "semPLS")
seminr_model <- estimate_model(data = mobi,
                               measurement_model = mobi_mm,
                               interaction = mobi_xm,
                               structural_model = mobi_sm)

print_paths(seminr_model)

seminr_model_boot <- bootstrap_model(data = mobi,
                                     measurement_model = mobi_mm,
                                     interaction = mobi_xm,
                                     structural_model = mobi_sm,
                                     nboot = 200)

print_paths(seminr_model_boot)
```

## Testing

To test:

```{r devtools-test, eval = FALSE}
require(devtools)
devtools::test()
```

