---
title: "SEMinR"
author: "Soumya Ray & Nicholas Danks"
date: "January 12, 2018"
output: rmarkdown::html_vignette 
vignette: >
  %\VignetteIndexEntry{SEMinR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# SEMinR

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(seminr)
```

There are now many packages available for R to build, estimate and make predictions
on PLS SEM models. However, the syntax of these models is disparate and complicated
and often discourages practitioners from use. Further, these packages require the 
manual computation and generation of interaction terms, do not extend bootstrap 
functionality to interaction models, and are not optimized to be used methodologically
in Montecarlo simulations and/or experiments.

The goal of SEMinR is to provide a natural syntax for researchers and methodologists 
to describe PLS structural equation models and that provides state of the art 
functionality and features - streamlined for high computational efficiency. Additionally,
the syntax is designed in such a way as to allow explicit definition of constructs as
either composite or common factor in nature - avoiding confusing use of formative in
the measurement model specification.

The description of the model often takes longer and requires dozens of lines of code,
however SEMinR seeks to automate this process and make modelling such features as
interactions and mediations much simpler.

There are 4 basic steps in the PLS SEM modelling:

* Describe the structural (inner) model

* Describe the measurement (outer) model

* Describe interactions

* Evaluate model performance

The SEMinR package makes these steps fast and easy:

* By providing easily readable and writable syntax, describing the model becomes simpler -
 and you can quickly change structural relationships allowing for multiple model comparisons.

* It provides simple "verbs", functions that correspond to the most common data manipulation tasks, to help you translate those thoughts into code.

* It automates such tedious tasks as creating interaction terms which usually require multiple lines of code and can be difficult to read and follow.

* By providing only the relevant and appropriate measurement-, structural- and overall-model evaluation criteria.  

This document introduces you to SEMinR's basic set of tools, and shows you how to apply them to your models. Other vignettes provide more details on specific topics:

* vignette#2: bla bla bla

* vignette#3: bla bla bla

## Data: 'mobi' 

To explore the basic syntax of SEMinR, we'll start with the data bundled with the seminr package - `mobi` data frame. The data set is used as measurement instrument for the european customer satisfaction index (ECSI) adapted to the mobile phone market, see Tenenhaus et al. (2005).

```{r}
dim(mobi)
head(mobi)
```

PLS SEM estimation and prediction packages generally take a source - target type matrix as parameters for each of the structural model and the measurement model. They require the practitioner to create any interaction terms by hand prior to model description. SEMinR makes use of simple and natural verbs to describe the model.

## Measurement model description verbs

SEMinR verbs for measurement model description:

* `multi_items()` or `single_item()`
* `composite()` or `reflective()`

These verbs are highly intuitive and provide for human-readable code that allows a practitioner to understand a measurement model on first glance and easily make changes to compare different measurement models. They also force the explicit definition of each construct
as belonging to either the composite or common-factor model.

### Describe multiple measurement items (manifest variables) with `multi_items()` or single item measurement items with `single_item()`

`single_item()` provides a verb to describe a single measurement item, while
`multi_items()` provides a verb to describe multiple measurement items in one simple line of human-readable code.

For example, we can describe the manifest variables: IMAG1, IMAG2, IMAG3, IMAG4 and IMAG5:

```{r, eval=FALSE}
multi_items("IMAG", 1:5)
```

`multi_items()` is used in conjunction with `composite()` or `reflective()` to describe a composite and common-factor construct respectively.

We can also describe a single manifest variable: CUSCO:

```{r, eval=FALSE}
single_item("CUSCO")
```

`multi_items()` is used in conjunction with `composite()` or `reflective()` to describe the multi-item measurement model of a composite and common-factor construct respectively.

`single_item()` is used in conjunction with composite()` or `reflective()` to describe the single-item measurement model of a composite and common-factor construct respectively. It is important to note that a single-item construct should be defined as either composite mode A or reflective.

### Describe a construct with `composite()` or `reflective()`

`composite()` and `reflective()` provide a verb to describe a construct in one simple line of human-readable code.

For example, we can use `composite()` to describe the mode A (or correlation weights) composite "Image" with manifest variables IMAG1, IMAG2, IMAG3, IMAG4 and IMAG5:

```{r, eval=FALSE}
composite("Image", multi_items("IMAG", 1:5), weights = "A")
```
or 
```{r, eval=FALSE}
composite("Image", multi_items("IMAG", 1:5), weights = "correlation")
```

or mode B (regression weights) composite "Image" with manifest variables IMAG1, IMAG2, IMAG3, IMAG4 and IMAG5:

```{r, eval=FALSE}
composite("Image", multi_items("IMAG", 1:5), weights = "B")
```
or
```{r, eval=FALSE}
composite("Image", multi_items("IMAG", 1:5), weights = "regression")
```

Similarly, we can use `reflective()` to describe the common-factor "Image" with manifest variables IMAG1, IMAG2, IMAG3, IMAG4 and IMAG5:

```{r, eval = FALSE}
reflective("Image", multi_items("IMAG", 1:5))
```

`multi_items()` is used in conjunction with `reflective()` or `composite()` to describe a construct.
`constructs()` is used in conjunction with multiple construct descriptions to describe the entire measurement model.

### Create measurement model matrix with `constructs()`

`constructs()` compiles the measurement model source-target matrix from the user specified construct descriptions described in the parameters.

For example, we can create a source-target matrix measurement model for the `mobi` data (NOTE: this is a demonstration model illustrating the options provided by seminr and does not reflect the model originally proposed by Tenenhaus et al., 2005).
Output is displayed:

```{r}
mobi_mm <- constructs(
  composite("Image",         multi_items("IMAG", 1:5), weights = "B"),
  composite("Expectation",   multi_items("CUEX", 1:3), weights = "regression"),
  composite("Quality",       multi_items("PERQ", 1:7), weights = "A"),
  composite("Value",         multi_items("PERV", 1:2), weights = "correlation"),
  reflective("Satisfaction", multi_items("CUSA", 1:3)),
  reflective("Complaints",   single_item("CUSCO")),
  reflective("Loyalty",      multi_items("CUSL", 1:3))
)
mobi_mm
```

The `seminr()` syntax is far more human-readable and the use of the verbs makes the definition of the measurement model explicit. Modifications to the `seminr()` measurement model description are quick, easy and intuitive.

## Structural model description verbs

SEMinR verbs for structural model description:

* `paths()`
* `relationships()`

These verbs are highly intuitive and provide for human-readable code that allows a practitioner to understand the structural model on first glance and easily make changes to compare different structural models.

### Describe structural model paths with `paths()`

`paths()` provides a verb to describe single or multiple structural paths.

For example, we can describe the structural model paths:

* "Image" -> "Expectation"
* "Image" -> "Satisfaction"
* "Value" -> "Satisfaction"

```{r, eval=FALSE}
paths(from = "Image",                     to = c("Expectation", "Satisfaction"))
paths(from = "Value",                     to = "Satisfaction")
paths(from = c("Satisfaction","Loyalty"), to = "Value")
```

`paths()` is used in conjunction with `relationships()` to describe the full structural model.

### Create structural model matrix with `relationships()`

`relationships()` compiles the structural model source-target matrix from the user specified structural path descriptions described in the parameters.

For example, we can create a source-target matrix structural model for the `mobi` data (with output displayed):

```{r}
mobi_sm <- relationships(
  paths(from = "Image",        to = c("Expectation", "Satisfaction", "Loyalty")),
  paths(from = "Expectation",  to = c("Quality", "Value", "Satisfaction")),
  paths(from = "Quality",      to = c("Value", "Satisfaction")),
  paths(from = "Value",        to = c("Satisfaction")),
  paths(from = "Satisfaction", to = c("Complaints", "Loyalty")),
  paths(from = "Complaints",   to = "Loyalty")
)
mobi_sm
```

Used in conjunction with the measurement model verbs, the description and comparison of models can is more intuitive, readable and comparable.

## Interaction term description verbs

SEMinR verbs for creating Interaction Terms:

* `interactions()`
* `interaction_ortho()` or `interaction_scaled()`

Most PLS SEM packages require practitioners to create interaction terms by hand which can be a time-consuming and arduous task - not to mention presenting many opporunities for errors and mis-specifications to enter the model. These verbs are highly intuitive and provide for human-readable code that allows a practitioner to quickly describe and "read" interaction terms and make changes and comparisons on the fly. 

Henseler and Chin (2010) suggest adjusting the SD of the interaction term because:
"In general, the product of two standardized variables does not equal the standardized product of these variables"
Seminr automatically adjusts for this providing highly accurate model estimations.

### Describe an interaction with `interaction_ortho()` or `interaction_scaled()`

`interaction_ortho()` provides a verb to describe single interaction item as generated by the orthogonalization method as described by Henseler and Chin (2010).
`interaction_scaled()` provides a verb to describe a single interaction item as generated by the scale product-indicator method as described by Henseler and Chin (2010).
For both these methods the SD of the interactino term is adjusted as noted above.

For example, we can describe the following interaction items:

* "Image" + "Expectation"
* "Image" + "Value"

Note: Seminr syntax uses a period as a naming convention for the interaction construct and thus "Image" + "Expectation" becomes "Image.Expectation". Please refrain from using a period in the naming of non-interaction constructs.

```{r, eval = FALSE}
interaction_ortho("Image", "Expectation")
interaction_scaled("Image", "Value")
```

These functions return functions themselves which are not resolved until passed as parameters to the `interactions()` function.

### Describe all model interactions with `interactions()`

`interactions()` provides a verb to describe all interactions in a specific model. The function itself returns a function which is not resolved until passed as a parameter to the `seminr()` function.

For example, we can describe the following interaction constructs in a model:

```{r}
mobi_xm <- interactions(
  interaction_ortho("Image", "Expectation"),
  interaction_ortho("Image", "Value")
)
```

The output of this code is:

```{r}
mobi_xm
```

As previously mentioned, a function which is not resolved until passed to `estimate_pls()`.

## PLS SEM Model Estimation with `seminr`

SEMinR verbs for PLS SEM model estimation:

* `estimate_pls()`

This verb is used to estimate the model and takes the following parameters:

* `data`: the dataset containing the measurement model items specified in `constructs()`
* `measurement_model`: the measurement model described by the `constructs()` function
* `interactions`: the interactions described by the `interactions()` function. If no interaction constructs are specified, this should be assigned the value of `NULL`
* `structural_model`: the structural model described by the `paths()` function
* `inner_weights`: the inner_weighting scheme for model estimation. This can take either `path_weighting` for path_weighting or `path_factorial` for factor weighting as proposed by Lohmöller (1989)

Note: `max_iterations` is set to 300 as default and `stop_criterion` to 7. 

For example, we can estimate the PLS SEM model based upon the structural and measurement model with interactions as described previously in this vignette.

```{r}
# seminr syntax for creating measurement model
mobi_mm <- constructs(
  reflective("Image",        multi_items("IMAG", 1:5)),
  reflective("Expectation",  multi_items("CUEX", 1:3)),
  reflective("Value",        multi_items("PERV", 1:2)),
  reflective("Satisfaction", multi_items("CUSA", 1:3))
)

# interaction factors must be created after the measurement model is defined
mobi_xm <- interactions(
  interaction_ortho("Image", "Expectation"),
  interaction_ortho("Image", "Value")
)

# structural model: note that name of the interactions factor should be
#  the names of its two main factors joined by a '.' in between.
mobi_sm <- relationships(
  paths(to = "Satisfaction",
        from = c("Image", "Expectation", "Value",
                 "Image.Expectation", "Image.Value"))
)

mobi_pls <- estimate_pls(data = mobi,
                         measurement_model = mobi_mm,
                         interactions = mobi_xm,
                         structural_model = mobi_sm,
                         inner_weights = path_weighting)
```

## Reporting the PLS SEM Model

There are multiple ways of reporting the estimated model. The `estimate_pls()` function returns an object of class `seminr_model`. This can be passed directly to the base R function `summary()`. This can be used in two primary ways:   

* `summary(seminr_model)` to report $R^{2}$, adjusted $R^{2}$, path coefficients for the structural model, and the construct reliability metrics $rho_{C}$, AVE, and $rho_{A}$.

* `model_summary <- summary(seminr_model)` returns an object of class `summary_seminr_model` which contains the following accessible objects:

** `model_summary$iterations` reports the number of iterations to converge on a stable model
** `model_summary$paths` reports the matrix of path coefficients, $R^{2}$, and adjusted $R^{2}$
** 

This verb is used to estimate the model and takes the following parameters:


The output of this code is:

```{r}
mobi_pls <- estimate_pls(data = mobi,
                         measurement_model = mobi_mm,
                         interactions = mobi_xm,
                         structural_model = mobi_sm,
                         inner_weights = path_weighting)
mobi_pls
report_paths(mobi_pls)
summary(mobi_pls)
```

## References

* Henseler, J., & Chin, W. W. (2010), A comparison of approaches for the analysis of interaction effects between latent variables 
using partial least squares path modeling. Structural Equation Modeling, 17(1), 82–109. https://doi.org/10.1080/10705510903439003
* Lohmöller, J.-B. (1989). Latent variables path modeling with partial least squares. Heidelberg, Germany: Physica- Verlag.
Marsh,
